A Hitchhiker's Guide to Reproducible Research" # replace with something serious

### 2021-04-30 manuscript.Rmd
small changes

Even when other researchers are excited to use your code or materials, they are usually legally inhibited from doing so.
In many countries (i.e. the USA, European Union), the creator of something has exclusive right to use their creation.
A license allows other people to use what you have created.
<!---exact copy from our previous manuscript, do we need to change? -->
In our experience, the [Creative Commons - Attribution license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media.
In contrast, scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT) (or similar permissive licenses).
Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.
These licenses are also in line with the recommendations by the Reproducible Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009].
A great resource to choose a license is [choosealicense.com](https://choosealicense.com).
However, no resource, including our recommendation, replaces legal advice.
<!-- plagiary end -->
It is important to get written consent from collaborators when you change/add the license.
However, if their first contribution occurs after adding a license, they give their consent by contributing.
To add a CC-BY license to your project, simply run (there are similar commands for other licenses):

@@

This command will go through all R Markdown files in your repository and infer a Makefile from them:

@@

Modern data analysis is only feasible because we rely on much external software.
One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility.
The critical problem is the fragility of a software setup.
Many components have to fit perfectly together, and each change, e.g. an update, risks that balance.
Updates are, of course, very much necessary.
Another problem is that the software environment is also variable and what works on one machine might not work on any other.
Docker circumvents such issues.
Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, down to the operating system.
Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
Such a snapshot of the software stack is called "image".
Of course, packaging all needed software in such an image requires considerable amounts of storage.
However, two strategies help to keep the storage requirements manageable.
First, there is a community that maintains images for particular purposes.
For example, using R [rocker-citation].
Such images only contain what is necessary for a specific purpose.
Users can then install whatever they additionally need on top of these purpose-built images.
The image that was used for this article uses `r round(as.numeric(fs::file_size("reprotutorial.sif"))/1024^3, 2)`Gb of disc space.
It does include R, RStudio, LaTeX as well as many R packages like the tidyverse (amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.)
Second, instead of saving a binary image of the software, one can also save a textual description of the programmatic steps to install the software in a `Dockerfile`.
Such files are tiny (the Dockerfile for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big), but they rely on the assumption that all software repositories are available and contain the correct software version.
The `repro`-package lets you specify required packages as part of the metadata of the RMarkdown:

@@

Docker can be a bit unfriendly, because it requires a complicated mental model and some knowledge about how to install software on Linux.
An R-focused alternative is the R package [`renv`](https://rstudio.github.io/renv/articles/renv.html)[^renv][@R-renv].
`renv` isolates all R packages used in a project and records there version as well as the source from where they where installed.
This mechanism is faster and simpler to use than Docker, at the expense of only tracking R packages.
`renv` therefore is already of great use for reproducibility.
However, even when your project relies solely on R, external dependencies still change the outcome of computations, like the operating system, the compilers used, the BLAS system in use and so forth.
Most importantly the R version itself has to match.

@@

One usecase in which Docker sometimes can't be used is in high performance computing (HPC).
The simulation we did is computationally expensive enough to warrant the use of a HPC cluster.
To run the simulation on a cluster we require the exact same software, not only on one or two computers, but potentially hundreds.
Docker is a natural solution for this problem and is the defacto standard for cloud computing.
However, Docker requires admin rights, while this is no issue on cloud computing infrastructure it is an issue in classical HPC environments.
Singularity was designed to run without admin rights, while maintaining full compatibility with Docker images.
If you already have a Docker image, you can choose if you want to run this image with Docker or with Singularity.
Anyway, from within the container i.e. for your analysis it does not make any difference.

@@

```{r}
# create a dummy data.frame with two columns
x <- data.frame(VAR1 = c(1, 2, 3, 4), VAR2 = c(0, 4, 6, 9))
# compute hash using md5
hash <- digest::digest(x, "md5")
if(hash != "4203854848ad0cc5d2f848c44b70815d"){
  warning("Mismatch between original and current simulations!\nHash now is:\n    '", hash, "'")
}
```

>>>>

Even when other researchers are excited to use your published code or materials, they are usually legally inhibited from doing so.
In many countries (i.e. the USA, European Union), the creator of something has exclusive right to use their creation.
A license allows other people to use what you have created.
<!---exact copy from our previous manuscript, do we need to change? -->
In our experience, the [Creative Commons - Attribution license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media.
In contrast, scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT) (or similar permissive licenses).
Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.
These licenses are also in line with the recommendations by the Reproducible Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009].
A great resource to choose a license is [choosealicense.com](https://choosealicense.com).
However, no resource, including our recommendation, replaces legal advice.
<!-- plagiary end -->
It is important to get written consent from collaborators when you change/add the license.
However, if their first contribution occurs after adding a license, they give their consent by contributing.
To add a CC-BY license to your project, simply run (there are similar commands for other licenses):

@@

This command will go through all R Markdown files, and their YAML metadata, in your repository and infer a Makefile from them:

@@

Modern data analysis is only feasible because we rely on much external software.
One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility.
The critical problem is the fragility of a software setup.
Many components have to fit perfectly together, and each change, e.g. an update, risks that balance.
Updates are, of course, very much necessary.
Another problem is that the software environment is also variable and what works on one machine might not work on any other.
Docker circumvents such issues.
Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, down to the operating system.
Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
Such a snapshot of the software stack is called "image".
Of course, packaging all needed software in such an image requires considerable amounts of storage.
However, two strategies help to keep the storage requirements manageable.
First, there is a community that maintains images for particular purposes.
For example, using R [rocker-citation].
Such images only contain what is necessary for a specific purpose.
Users can then install whatever they additionally need on top of these purpose-built images.
The image that was used for this article uses `r round(as.numeric(fs::file_size("reprotutorial.sif"))/1024^3, 2)`Gb of disc space.
It does include Ubuntu (a Linux operating system) R, RStudio, LaTeX as well as many R packages like the tidyverse (amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.)
Second, instead of saving a binary image of the software, one can also save a textual description of the programmatic steps to install the software in a `Dockerfile`.
Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big), but they rely on the assumption that all software repositories are available and contain the correct software version.
The `repro`-package lets you specify required packages as part of the metadata of the RMarkdown:

@@

Docker can be a bit unfriendly, because it requires a complicated mental model and some knowledge about how to install software on Linux.
An R-focused alternative is the R package [`renv`](https://rstudio.github.io/renv/articles/renv.html)[^renv][@R-renv].
`renv` isolates all R packages used in a project and records there version as well as the source from where they where installed.
This mechanism is faster and simpler to use than Docker, at the expense of only tracking R packages.
`renv` therefore is already of great use for reproducibility.
However, even when your project relies solely on R, external dependencies can still change the outcome of computations, like the operating system, the compilers used, the BLAS system in use and so forth.
Most importantly the R version itself has to match.

@@

One usecase in which Docker sometimes can not be used is in high performance computing (HPC).
The simulation we did is computationally expensive enough to warrant the use of a HPC cluster.
To run the simulation on a cluster we require the exact same software, not only on one or two computers, but potentially hundreds.
Docker is a natural solution for this problem and is the defacto standard for cloud computing.
However, Docker requires admin rights, while this is no issue on cloud computing infrastructure it is an issue in classical HPC environments.
Singularity was designed to run without admin rights, while maintaining full compatibility with Docker images.
If you already have a Docker image, you can choose if you want to run this image with Docker or with Singularity.
Anyway, from within the container i.e. for your analysis it does not make any difference.

@@

```{r}
# create a dummy data.frame with two columns
x <- data.frame(VAR1 = c(1, 2, 3, 4), VAR2 = c(0, 4, 6, 9))
# compute hash using md5
hash <- digest::digest(x, "md5")
if(hash != "5ba412f5a26f43842971dd74954fcdeb"){
  warning("Mismatch between original and current data!\nHash now is:\n    '", hash, "'")
}
```
